Virtual Memory Assisted Transactional Memory for Dynamic Languages

### 2014-05-27 talk/dls2014/paper/paper.tex
re-run with no-jit build and add scaling micro-bench

The \emph{page privatisation}, which represents the ratio between
private pages and shared pages, gives us an idea about how many
private copies exist for each shared page. If the ratio is $>1.0$, it
means that on average we have more than one private copy for each
shared page. In the worst case described above, this number would
reach $N$ and means we need $N\times$ the memory for the private
pages alone. In this benchmark, we see the same spikes as the memory
usage. These come directly from re-sharing the pages. The maximum page
privatisation is around $0.5$ between major collections, thus the
private pages are responsible for a $50\%$ increase in the required
memory. Since the spikes in the GC managed memory line actually
show increases by $~80\%$, it means that the rest comes from actual
garbage objects that were collected.

For PyPy-STM the average memory requirement is 29~MiB and there are
$\sim 11$ major collections during the runtime. Normal PyPy with a GIL
grows its memory up to just 7~MiB and does not do a single major
collection in that time. Compared to normal PyPy, we are missing a
memory optimisation to store small objects in a more compact
way. Additionally, since normal PyPy uses a GIL, it does not need to
duplicate any data structures like e.g. the Nursery for each
thread. This, the missing optimisation, and the additional memory
requirements for STM explained above account for this difference.
\remi{I don't know how much sense it makes to go deeper. We will
improve this in the future, but right now this is the overall picture.}

@@

maybe some simple micro benchmarks with adaptable conflict rate

>>>>

The \emph{page privatisation}, which represents the ratio between
private pages and shared pages, gives us an idea about how many
private copies exist for each shared page. If the ratio is $>1.0$, it
means that on average we have more than one private copy for each
shared page. In the worst case described above, this number would
reach $N$ and means we need $N\times$ the memory for the private
pages alone. In this benchmark, we see the same spikes as the memory
usage. These come directly from re-sharing the pages. The maximum page
privatisation is around $0.9$ between major collections, thus the
private pages are responsible for a $90\%$ increase in the required
memory. Since the spikes in the GC managed memory line actually
show increases by $~100\%$, it means that the rest comes from actual
garbage objects that were collected.

For PyPy-STM the average memory requirement is 21~MiB and there are
$\sim 10$ major collections during the runtime. Normal PyPy with a GIL
grows its memory up to just 7~MiB and does not do a single major
collection in that time. Compared to normal PyPy, we are missing a
memory optimisation to store small objects in a more compact
way. Additionally, since normal PyPy uses a GIL, it does not need to
duplicate any data structures like e.g. the Nursery for each
thread. This, the missing optimisation, and the additional memory
requirements for STM explained above account for this difference.
We expect to improve this aspect in the future, in this paper we
want to focus first on performance.
\remi{I don't know how much sense it makes to go deeper. We will
improve this in the future, but right now this is the overall picture.}

@@

To asses how well the STM system scales on its own (without any real
workload), we execute the following loop on 1 to 4 threads:
\begin{lstlisting}
def workload():
    i = 20000000
    while i:
        i -= 1
\end{lstlisting}

For the results in figure \ref{fig:scaling}, we averaged
over 5 runs and normalised the average runtimes to the
time it took on a single thread. From this we see that there
is additional overhead introduced by each thread ($13\%$
for all 4 threads together).

\remi{what we don't show is by how much this overhead is influenced
by allocations}

\begin{figure}[h]
  \centering
  \includegraphics[width=1\columnwidth]{plots/scaling.pdf}
  \caption{Scalability of the STM system\label{fig:scaling}}
\end{figure}
